---
title: "Heart Disease Prediction"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Heart-Disease-Prediction
Analyzing a data set to predict the presence of any heart disease in a patient 
######## Attaching file and giving names to its columns(variables) #######


## Required libraries
```{r}
#install.packages('class')
#install.packages('gmodels')
#install.packages('caret')
#install.packages('e1071')
#install.packages('corrplot')
#install.packages('sqldf')
library('class')
library('gmodels')
library('caret')
library('e1071')
library('corrplot')
library('sqldf')
```

## Including Code

```{r}
heart<-read.csv("C:/Users/17815/Documents/Heart-Disease-Prediction-master/Heart Disease dataset.csv", header = FALSE)
head(heart)
names(heart)<-c("age","sex","chest_pain","resting_bp","cholestrol","fasting_blood_sugar","resting_ecg","max_heart_rate","exercise_induced_angina","oldpeak","slope_of_peak_exercise","number of major vessels colored","thallium heart scan","result")
#View(heart)
sapply(heart, class)
```

## Detecting and removing NA values


```{r}
heart[heart=="?"]<-NA
nrow(heart[is.na(heart$`number of major vessels colored`) | is.na(heart$`thallium heart scan`),])
nrow(heart)
data <- heart[!(is.na(heart$`number of major vessels colored`) | is.na(heart$`thallium heart scan`)),]
nrow(data)
```

```{r}
########## scatter-plot matrix ##########

pairs(~chest_pain +  max_heart_rate +  result, data=data)


#######Correlation matrix ############
data1<-sqldf("SELECT age as Age, sex as Sex, chest_pain as CP, resting_bp as RBP, 
             cholestrol as Cholestrol, fasting_blood_sugar as FBS, resting_ecg as ECG,
             max_heart_rate as HR, exercise_induced_angina as Angina, result as res FROM data")

corMatrix <- cor(data1)

########Correlation matrix##########

par(mfrow=c(1,1))
corrplot(corMatrix, method="circle", type="lower", addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=90, #Text label color and rotation
         diag=TRUE, sig.level = 0.05, insig = "blank")

```


## Converting to numeric form 

```{r}
data$age <- as.numeric(data$age)
data$`chest_pain` <- as.numeric(data$`chest_pain`)
data$`fasting_blood_sugar` <- as.numeric(data$`fasting_blood_sugar`)
data$`resting_ecg` <- as.numeric(data$`resting_ecg`)
data$`exercise_induced_angina` <- as.numeric(data$`exercise_induced_angina`)
data$`slope_of_peak_exercise` <- as.numeric(data$`slope_of_peak_exercise`)
data$`resting_bp` <- as.numeric(data$`resting_bp`)
data$`thallium heart scan` <- as.numeric(data$`thallium heart scan`)
data$cholestrol <- as.numeric(data$cholestrol)
data$sex <- as.numeric(data$sex)
data$`number of major vessels colored` <- as.numeric(data$`number of major vessels colored`)
```

## Normalizing the dataset

```{r}
normalize <- function(x) {
        return((x - min(x)) / (max(x) - min(x)))
}

data_n <- as.data.frame(lapply(data[1:13], normalize))
summary(data_n)
```

## Splitting the dataset in to train & test

```{r}
set.seed(1000)

# random selection of 70% of data       
rand.70 <- sample(1:nrow(data_n),size=nrow(data_n)*0.7,replace = FALSE)

# Training set
train_set <- data_n[rand.70,]   # 70% training data
test_set <- data_n[-rand.70,]   # 30% test data

# Target set
# Creating a data frame for 'defaulter' feature which is our result
train_target <- data[rand.70,14]
test_target <- as.factor(data[-rand.70,14])

```

## KNN Classification
```{r}
sqrt(297) # total observations are 297
knn.17 <- as.factor(knn(train = train_set, test = test_set, cl = train_target, k = 17))
table(knn.17, test_target)
ACC.173 <- 100 * sum(test_target == knn.17)/NROW(test_target)
ACC.173  

# obtaining the cross table
CrossTable(x = test_target, y = knn.17, prop.chisq = FALSE)
confusionMatrix(test_target , knn.17)
```

## Logistic Regression
```{r}
female<-subset(data,sex==0)
male<-subset(data,sex=1)
length(female)

#For females

logistic_female <- glm(result ~ ., data=female, family="binomial")
summary(logistic_female)
pred_f <- predict(logistic_female, female[-14],  type = "response")
female$prediction <- ifelse(pred_f >= 0.5, "1", "0")
female$prediction <- as.factor(female$prediction)
female$result <- as.factor(female$result)
sapply(female, class)
confusionMatrix(female$result , female$prediction)

#For males

logistic_male<-glm(result~., data=male, family= "binomial")
summary(logistic_male)
pred_m <- predict(logistic_male, male[-14],  type = "response")
male$prediction <- ifelse(pred_m >= 0.5, "1", "0")
male$prediction <- as.factor(male$prediction)
male$result <- as.factor(male$result)
confusionMatrix(male$result , male$prediction)
```

